---
title: "Site selection for postfire aspen project"
author: "Sarah Carter and Sarah Hart"
email: "Sarah.Hart@colostate.edu"
date: "June 16, 2022"
output: 
  bookdown::html_document2:
    code_folding: hide
    toc: TRUE
    toc_depth: 4
    toc_float: TRUE
    collapsed: TRUE
    theme: spacelab
    fig_caption: TRUE
    number_sections: FALSE
citationcolor: blue
csl: "`r here:::here('ecology.csl')`"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, dpi = 600, progress = FALSE, cache = FALSE, fig.width=5,  fig.height=3.5, fig.align="center")

set.seed(513)
options(repos = c(CRAN = "http://cran.rstudio.com"))
options(timeout=60*60*2) #timeout downloads that last longer than an hour
options(na.action = "na.fail")  

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  here, # easy file structure
  tidyverse, # data manipulation
  tidymodels, 
  ggplot2, # figures
  patchwork, # multiple figures
  readxl, # read excel files
  ggpubr, # plotting
  ggridges, # plotting
  lme4, # generalized linear mixed effects models
  effects, # plotting effects from glmms
  DHARMa # evaluating residuals from glmms
) 


# Set custom plotting theme
theme_new <- function(base_size = 9,base_family = "Helvetica"){
  theme_classic(base_size = base_size, base_family = base_family) %+replace%
    theme(
      axis.line.x = element_line(color="black", size = 0.25),
      axis.line.y = element_line(color="black", size = 0.25),
      axis.title = element_text(size = 9),
      axis.text = element_text(colour="black", size=8),
      legend.key=element_rect(colour=NA, fill =NA),
      panel.grid = element_blank(),   
      plot.background = element_rect(fill = NA, colour = NA),
      panel.border = element_rect(fill = NA, colour = NA),
      panel.background = element_rect(fill = "white", colour = "black"), 
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(size = 9)
      
    )
}
theme_set(theme_new())

options(ggplot2.continuous.colour="viridis")
options(scipen=999)

# Set directory structure for project
dir.create(here("Data/"), showWarnings = FALSE)
dir.create(here("Results/"), showWarnings = FALSE)
```
# Overview
## Research questions
(1) Where do aspen seedlings establish?
(2) What landscape factors explain aspen seedling density?
(3) What landscape and microsite factors, including competition with suckers, influence aspen seedling occurrence?
(4) What landscape and microsite factors, including competition with suckers, influence aspen seedling height?

## Hypotheses

# Methods
## Study area

## Field methods
## Data analysis
### Import data
```{r importData}
POTR <- read_excel(here("Data", "POTR Seedling Data-2.xlsx"))
#str(POTR)
colnames(POTR) <- tolower(c("SITE.NAME", "SITE.NO", "TRANSECT", "SUBPLOT", "HEIGHT", "SUBSTRATE", "SM.TOPO", "LG.TOPO", "LG.CWD", "SM.CWD","SUCKER.DIST", "CC", "BROWSE"))

POTR_site <- read_excel(here("Data", "POTR SITE DATA.xlsx"))
#str(POTR_site)
colnames(POTR_site) <- tolower(c("SITE.NO", "CLUSTER", "UTM.E", "UTM.N", "ELEVATION", "SLOPE", "ASPECT", "TOPO", "TRANS.A.DIR", "TRANS.B.DIR", "LIVE.DIST", "DEAD.DIST"))

REGEN <- read_excel(here("Data", "ALL REGEN.xlsx"))
#str(REGEN)
colnames(REGEN) <- tolower(c("SITE.NO", "TRANSECT", "POINT", "ABLA", "PICO", "PIEN", "PIPO", "PSME", "POTR.sucker", "POTR.seed"))

#rename variables because they are confusing 
#POTR$site.name <- POTR$SITE.NAME
#POTR$site.Number <- POTR$SITE..
#POTR$height <- POTR$Height..cm.
```

### Data wrangling
```{r}
compiled <- merge(POTR, POTR_site, by = "site.no")
#compiled2 <- REGEN %>% group_by(site.no) %>% summarize_at(.vars=vars(abla:potr.seed), .fun=sum)
compiled.REGEN <- merge(REGEN, 
POTR_site,  by = c("site.no"))
compiled.REGEN$potr.seed01 <- ifelse(compiled.REGEN $potr.seed>0, 1,0)
```

### Notes 
- what are negative values of slope?


### Where do aspen seedlings establish?

To understand how site conditions influence seedling recruitment, we
first modeled the presence/absence of successful seedling germination
using a generalized linear mixed effect model (GLMM) with a
binomial-distributed error structure, a logit link, and random
intercepts of site identity. Next, to understand how site conditions
influence the count of seedlings present at each block, we modeled the
total number of seedlings that germinated at block where successful
germination occurred using a GLMM. To account for the high variability
in the count of seedlings per block, we used a negative binomial GLMM
and again included a random intercept of site identity.

```{r}

occurrence.mod.cluster<- glmer(potr.seed01 ~ scale(elevation) + scale(aspect)+ (1|cluster), family=binomial, data=compiled.REGEN)
occurrence.mod.cluster.site<- glmer(potr.seed01 ~ scale(elevation) + (1|cluster) + (1|site.no:cluster), family=binomial, data=compiled.REGEN)
occurrence.mod.cluster.site.transect<- glmer(potr.seed01 ~ scale(elevation) + scale(slope) + scale(aspect) + (1|cluster) + (1|site.no:cluster) + (1|transect:site.no:cluster), family=binomial, data=compiled.REGEN)

occurrence.mod.cluster%>% AIC()
occurrence.mod.cluster.site%>% AIC()
occurrence.mod.cluster.site.transect%>% AIC()

occurrence.mod.cluster.site.transect%>% predictorEffects(c('elevation', 'slope', 'aspect'), xlevels=100) %>% plot()
e.moist<- predictorEffect("elevation",occurrence.mod, xlevels=100) %>% as.data.frame()

library(DHARMa)
modfinal.binomial.simout <- simulateResiduals(occurrence.mod.cluster.site.transect)
plot(modfinal.binomial.simout)
```


```{r}
ggplot(compiled2, aes(y=elevation, x=factor(potr.seed01)))+geom_boxplot()

ggplot(compiled2, aes(y=slope, x=factor(potr.seed01)))+geom_boxplot()
```

## Seedling height by cluster
```{r}
#dotplot height by cluster
ggplot(compiled,aes(x=site.name,y=height, col = height)) + geom_point() + 
    xlab("Site Cluster Name") + 
    ylab(expression(paste("Height (cm)"))) +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle(" Seedling Height by Cluster")
```
```{r}
#boxplot height by cluster
ggplot(POTR, aes(x=site.name, y=height)) + 
  geom_boxplot() +
  xlab("Site Cluster") +
  ylab(expression(paste("Height (cm)"))) +
  ggtitle("Seedling Height by Cluster")

```
```{r}
#histogram counts
count <- count(compiled, site.name)
count

#filter for seedling density
dense <- compiled %>% filter(site.name == "CAM"| site.name == "LONG"| site.name == "MONTY"| site.name == "RAWAH" | site.name == "SNOW")

#height by cluster for dense clusters 
ggplot(dense,aes(x=site.name,y=height, col = height)) + geom_point() + 
    xlab("Site Cluster Name") + 
    ylab(expression(paste("Height cm"))) +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle(" Seedling Height by Cluster")

```
```{r}

ggline(dense, x = "site.name", y = "height", color = "site.name", add = c("mean_se", "jitter"),  ylab = "Height (cm)", xlab = "Cluster", labcol = "CLuster")

```
```{r}
tapply(compiled$height, compiled$site.name, summary)
```
```{r}
ggplot(POTR, aes(x = height, y = site.name, fill=site.name)) +
  geom_density_ridges() +
  xlab("Height cm") +
  ylab(expression(paste("Cluster Name"))) +
  ggtitle("Seedling Height Distribution")
```

#data filtering for tests
```{r}
library(dplyr)

#by cluster only
snow <- compiled %>% filter(SITE.NAME =="SNOW")
res <- compiled %>% filter(SITE.NAME =="RES")
rawah <- compiled %>% filter(SITE.NAME =="RAWAH")
monty <- compiled %>% filter(SITE.NAME =="MONTY")
long <- compiled %>% filter(SITE.NAME =="LONG")
lake <- compiled %>% filter(SITE.NAME =="LAKE")
fish <- compiled %>% filter(SITE.NAME =="FISH")
elkhorn <- compiled %>% filter(SITE.NAME =="ELKHORN")
cr69 <- compiled %>% filter(SITE.NAME =="CR69")
cam <- compiled %>% filter(SITE.NAME =="CAM")
blue <- compiled %>% filter(SITE.NAME =="BLUE")

#by substrate
ash <- compiled %>% filter(Substrate == "A"| Substrate == "A/B"| Substrate == "A/L"| Substrate == "A/M")
bryophyte <- compiled %>% filter(Substrate == "B"| Substrate == "B/A"| Substrate == "B/L"| Substrate == "B/M")
litter <- compiled %>% filter(Substrate == "L"| Substrate == "L/A"| Substrate == "L/B"| Substrate == "L/M")
mineral <- compiled %>% filter(Substrate == "M"| Substrate == "M/A"| Substrate == "M/B"| Substrate == "M/L")
rock <- compiled %>% filter(Substrate == "R")
wood <- compiled %>% filter(Substrate == "W")

#by small topo
cc <- compiled %>% filter(Small.Topo == "CC")
cv <- compiled %>% filter(Small.Topo == "CV")
f <- compiled %>% filter(Small.Topo == "F")
s <- compiled %>% filter(Small.Topo == "S")

#by large topo
CC <- compiled %>% filter(Large.Topo == "CC")
CV <- compiled %>% filter(Large.Topo == "CV")
F <- compiled %>% filter(Large.Topo == "F")
S <- compiled %>% filter(Large.Topo == "S")

#by elevation groups
high <- compiled %>% filter(Elevation > "3064")
moderate <- compiled %>% filter(between(Elevation, 2713,3064))
low <- compiled %>% filter(Elevation < "2713") 

#by large CWD
Lcwdp <- compiled %>% filter(Large.CWD == "1")
Lcwda <- compiled %>% filter(Large.CWD == "0")

#by small CWD
Scwdp <- compiled %>% filter(Small.CWD == "1")
Scwda <- compiled %>% filter(Small.CWD == "0")
```

#seedling count variable
```{r}
seedling_count_site <- count(compiled, compiled$SITE..)
seedling_count_cluster <- count(compiled, compiled$SITE.NAME)
```

#method of deciding relative site elevation groups
```{r}
#median(POTR_site$Elevation)
#min(POTR_site$Elevation)
#max(POTR_site$Elevation)
#summary(POTR_site$Elevation)
```



#substrate tallies and chi-sq
```{r}
count(ash)
count(bryophyte)
count(litter)
count(mineral)
count(rock)
count(wood)

substrate_counts = c(224, 256, 8, 55, 0, 4)
#chisq.test(substrate_counts, correct = F)

```

#topo
```{r}
cc
cv
s
f
CC
CV
S
F
```


#relationships between height and other variables
```{r}
#height and substrate
kruskal.test(Height..cm. ~ Substrate, data = compiled)

library(FSA)
dunnTest(Height..cm. ~ Substrate, data = compiled, method = "holm")

#height and cluster
kruskal.test(Height..cm. ~ SITE.NAME, data = compiled)

dunnTest(Height..cm. ~ SITE.NAME, data = compiled, method = "holm")

#height and elevation

kruskal.test(Height..cm. ~ Elevation, data = compiled)

dunnTest(Height..cm. ~ Elevation, data = compiled, method = "holm")

#small topo and height

kruskal.test(Height..cm. ~ Small.Topo, data = compiled)

dunnTest(Height..cm. ~ Small.Topo, data = compiled, method = "holm")

#large topo and height

kruskal.test(Height..cm. ~ Large.Topo, data = compiled)

dunnTest(Height..cm. ~ Large.Topo, data = compiled, method = "holm")


```

#Kruskall tests with seedling count
```{r}
#count and cluster
kruskal.test(seedling ~ SITE.NAME, data = compiled)

dunnTest(seedling ~ SITE.NAME, data = compiled, method = "holm")

#count and elevation

kruskal.test(seedling ~ Elevation, data = compiled)

dunnTest(seedling ~ Elevation, data = compiled, method = "holm")

#count and topo 

kruskal.test(seedling ~ Small.Topo, data = compiled)

dunnTest(seedling ~ Small.Topo, data = compiled, method = "holm")


#count topo and height

kruskal.test(seedling ~ Large.Topo, data = compiled)

dunnTest(seedling ~ Large.Topo, data = compiled, method = "holm")
```


#substrate
```{r}
ash
bryophyte
litter
mineral
rock
wood
```




#elevation hist
```{r}
low
moderate
high
hist(compiled$Elevation, breaks = 10, col = "light blue", main = "Elevation range of seedlings across all sites", xlab = "Elevation (m)", ylab = "Frequency")
```
#browse and CWD

```{r}
#by large CWD
Lcwdp 
Lcwda 

#by small CWD
Scwdp 
Scwda 
```

#all regen
```{r}
#REGEN
abla <- REGEN %>% filter(ABLA > "0")
pico <- REGEN %>% filter(PICO > "0")
pien <- REGEN %>% filter(PIEN > "0")
pipo <- REGEN %>% filter(PIPO > "0")
psme <- REGEN %>% filter(PSME > "0")

sucker_regen <- REGEN %>% filter(POTR..sucker. > "0")

abla
pico
pien
pipo
psme
sucker_regen
```

#other aspen
```{r}
sucker <- compiled %>% filter(Sucker.Dist. < "51")
parent <- compiled %>% filter(Distance.to.nearest.live.aspen < "51")
orphan <- compiled %>% filter(Distance.to.nearest.dead.aspen < "51")

parent
orphan

```
#browse
```{r}
browse <- compiled %>% filter(Browse == "1")
no_browse <- compiled %>% filter(Browse == "0")

browse
tapply(browse$Height..cm., browse$SITE.., summary)
```

```{r}
compiled
seedling_count_cluster
```

```{r}
TransectCount <- count(compiled, compiled$Transect)
```

#how to set up binomial linear mixed model??
```{r}
n <- 38
id <- seq(n)
transect <- 1:2
d <- expand.grid(id = id, transect=transect)
set.seed(1)
#sect <- sample(c("A", "B"), size = n, replace = TRUE)
cluster <- sample(c("Blue", "Cam", "CR69", "Elkhorn", "Fish", "Lake", "Long", "Monty", "Rawah", "Res", "Snow"), size = n, replace = TRUE)
d$cluster <- cluster[d$id]
#d$sect <- sect[d$id]
d <- d[order(d$id, d$transect),]
rownames(d) <- NULL
head(d, n = 76)
```

Probabilities of seedling presence:
```{r}
#Blue p = 
#Cam
#CR
#Elk
#Fish
#Lake
#Long
#Monty
#Rawah
#Res
#Snow



d$clusterint <- interaction(d$cluster)
probs <- c()
```

#Visualizing abnormality
```{r}
#hist(seedling_count_cluster$n)
#hist(seedling_count$n)
#hist(compiled$Height..cm.)
```





















