---
title: "Site selection for postfire aspen project"
author: "Sarah Carter and Sarah Hart"
email: "Sarah.Hart@colostate.edu"
date: "June 16, 2022"
output: 
  bookdown::html_document2:
    code_folding: hide
    toc: TRUE
    toc_depth: 4
    toc_float: TRUE
    collapsed: TRUE
    theme: spacelab
    fig_caption: TRUE
    number_sections: FALSE
citationcolor: blue
csl: "`r here:::here('ecology.csl')`"
bibliography: "`r here:::here('MyLib.bib')`"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, dpi = 600, progress = FALSE, cache = FALSE, fig.width=5,  fig.height=3.5, fig.align="center")

set.seed(513)
options(repos = c(CRAN = "http://cran.rstudio.com"))
options(timeout=60*60*2) #timeout downloads that last longer than an hour
options(na.action = "na.fail")  

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  here, # easy file structure
  tidyverse, # data manipulation
  tidymodels, 
  reshape, # old data manipulation pkg
  ggplot2, # figures
  patchwork, # multiple figures
  grid, # multiple figures
  ggsci, # color palettes
  colorspace, # color palettes
  RColorBrewer, # color palettes
  knitr, # markdown documents
  kableExtra, # tables in markdown
  flextable, # other table options
  bookdown, # figure numbering in markdown
  sp, # spatial data (old pkg)
  sf, # spatial data (new pkg)
  raster, # raster data (old pkg)
  fasterize, # quick polygon to raster
  stars, # raster data (new pkg)
  starsExtra, # extra raster functions
  tigris, # US Census bureau GIS data
  tmap # easy plotting of maps
) 


# Set custom plotting theme
theme_new <- function(base_size = 9,base_family = "Helvetica"){
  theme_classic(base_size = base_size, base_family = base_family) %+replace%
    theme(
      axis.line.x = element_line(color="black", size = 0.25),
      axis.line.y = element_line(color="black", size = 0.25),
      axis.title = element_text(size = 9),
      axis.text = element_text(colour="black", size=8),
      legend.key=element_rect(colour=NA, fill =NA),
      panel.grid = element_blank(),   
      plot.background = element_rect(fill = NA, colour = NA),
      panel.border = element_rect(fill = NA, colour = NA),
      panel.background = element_rect(fill = "white", colour = "black"), 
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(size = 9)
      
    )
}
theme_set(theme_new())

options(ggplot2.continuous.colour="viridis")
options(scipen=999)

# Set directory structure for project
dir.create(here("Data/"), showWarnings = FALSE)
dir.create(here("Results/"), showWarnings = FALSE)
```

# Objective

The purpose of this code is to identify accessible sites that burned at high severity within the Cameron Peak Fire that were nearby existing patches of aspen.

## Data
To identify potential sites. We first obtained GIS data describing the extent of the Cameron Peak fire from the Monitoring Trends in Burn Severity Project,

```{r, eval=F}
dir.create(here("Data/Spatial/MTBS"), showWarnings = FALSE)
temp <- here("Data/Spatial/MTBS/MTBS.zip")
download.file("https://edcintl.cr.usgs.gov/downloads/sciweb1/shared/MTBS_Fire/data/composite_data/burned_area_extent_shapefile/mtbs_perimeter_data.zip", temp)
unzip(temp, exdir=here("Data/Spatial/MTBS/"))
```

We also obtained spatially data describing land ownership from the USFS and the Colorado Natural Heritage Program and the Geospatial Centroid.

```{r, eval=F}
dir.create(here("Data/Spatial/USFS"), showWarnings = FALSE)
temp <- here(paste0("Data/Spatial/USFS/RangerDistricts.zip"))
download.file("https://data.fs.usda.gov/geodata/edw/edw_resources/shp/S_USA.RangerDistrict.zip", temp)
unzip(temp, exdir=here("Data/Spatial/USFS/"))
```

```{r, eval=F}
dir.create(here("Data/Spatial/CoMaP"), showWarnings = FALSE)
temp <- here(paste0("Data/Spatial/CoMaP/CoMaP_GDB.zip"))
download.file("https://comap.cnhp.colostate.edu/dev/mapcollab/site/COMaP_GDB", temp)
unzip(temp, exdir=here("Data/Spatial/CoMaP"))
```

To identify sites that were easily accessible, we acquired spatial data describing the location of roads from CDOT and the USFS and the location of trails from the USFS.

```{r  download CDOT roads data, eval=F}
dir.create(here("Data/Spatial/Roads"), showWarnings = FALSE)
temp <- here("Data/Spatial/Roads/CO_HWY.zip")
download.file("http://dtdapps.coloradodot.info/staticdata/Downloads/StateGeoData/highways.zip", temp )
unzip(temp, exdir=here("Data/Spatial/Roads/"))

temp <- here("Data/Spatial/Roads/CO_LocalRoads.zip")
download.file("http://dtdapps.coloradodot.info/staticdata/Downloads/StateGeoData/LocalRoads.zip", temp )
unzip(temp, exdir=here("Data/Spatial/Roads/"))

temp <- here("Data/Spatial/Roads/CO_MajorRoads.zip")
download.file("http://dtdapps.coloradodot.info/staticdata/Downloads/StateGeoData/MajorRoads.zip", temp )
unzip(temp, exdir=here("Data/Spatial/Roads/"))
```

```{r download USFS roads data, eval=F}
dir.create(here("Data/Spatial/Roads"), showWarnings = FALSE)
temp <- here("Data/Spatial/Roads/S_USA.Road_MVUM.zip")
download.file("https://data.fs.usda.gov/geodata/edw/edw_resources/shp/S_USA.Road_MVUM.zip", temp )
unzip(temp, exdir=here("Data/Spatial/Roads/"))
```

```{r download trails data, eval=F}
dir.create(here("Data/Spatial/Trails"), showWarnings = FALSE)
temp <- here("Data/Spatial/Roads/S_USA.TrailNFS_Publish.zip")
download.file("https://data.fs.usda.gov/geodata/edw/edw_resources/fc/S_USA.TrailNFS_Publish.gdb.zip", temp )
unzip(temp, exdir=here("Data/Spatial/Trails/"))
```

To quantify the proximity of sites to pre-fire aspen cover, we acquired a 10 m resolution map of aspen cover for the Southern Rocky Mountain Ecoregion from Cook et al. (in review). The map represents aspen cover in ca. 2019 and was produced from Sentinel imagery.

```{r, eval=F}
dir.create(here("Data/Spatial/Aspen"), showWarnings = FALSE)
temp <- here("Data/Spatial/Aspen/SRM_ASPEN_CLASSIFIED_RF_10M_V1_0522.tif")
download.file("https://drive.google.com/file/d/1uhV4rCJ6DM26HoyNU-emuh3rPxrR807Q/view?usp=sharing", temp )
```

## Identification of potential study sites

1. Select areas of public land that burned at high severity
```{r, eval=F}
fireseverity <-  read_stars(here("Data/Spatial/MTBS/composite_data/MTBS_BSmosaics/2020/mtbs_CONUS_2020/mtbs_CONUS_2020.tif"))

select.fires <- st_read(here("Data/Spatial/MTBS/mtbs_perims_DD.shp")) %>% 
  filter(Incid_Name == "CAMERON PEAK") %>%
  st_transform(st_crs(fireseverity))  %>% 
  st_buffer(dist=-50)

comap <- st_read(here("Data/Spatial/CoMaP/COMAP-public.shp"))  %>%
  st_transform(st_crs(fireseverity))  # note for some reason I couldn't read the CoMAP data into R directly, so this datashet was generated by selecting all public land in QGIS
burned.public <- st_intersection(comap, select.fires)  %>% st_union()

st_write(burned.public, here("Data/Spatial/Sampling/Burned-public.shp"), append=F)
```

2. Select roads that are easy enough to drive

```{r, eval=F}
select.usfs.roads <- st_read(here("Data/Spatial/Roads/S_USA.Road_MVUM.shp")) %>% 
  filter(OPERATIONA %in% c("5 - HIGH DEGREE OF USER COMFORT", "4 - MODERATE DEGREE OF USER COMFORT", "3 - SUITABLE FOR PASSENGER CARS"))  %>%
  st_transform(st_crs(select.fires)) %>% # change projection
  st_crop(st_bbox(select.fires)) %>% # crop to study area
  st_union() 
```

```{r, eval=F}
hwys <- rgdal::readOGR(here("Data/Spatial/Roads/SHP/STATEWIDE/Highways.shp")) %>% # note for some reason the sf package could not read in this file but rgdal could
  st_as_sf() %>% 
  st_zm() %>%
  st_transform(st_crs(select.fires)) %>% 
  st_crop(st_bbox(select.fires)) %>%  # crop to study area
  st_union()
local <- st_read(here("Data/Spatial/Roads/SHP/STATEWIDE/local_roads.shp")) %>%
  st_zm() %>%
  st_transform(crs=st_crs(select.fires))  %>% 
  st_crop(st_bbox(select.fires)) %>%  # crop to study area
  st_union()
major <- st_read(here("Data/Spatial/Roads/SHP/STATEWIDE/Major_Roads.shp")) %>%
  st_zm() %>% 
  st_transform(crs=st_crs(select.fires)) %>% 
  st_crop(st_bbox(select.fires)) %>%  # crop to study area
  st_union()
```

3. Select trails that intersect the fire

```{r, eval=F}
trails <- st_read(here("Data/Spatial/Trails/Trails.shp")) %>%
  st_transform(st_crs(select.fires)) %>% 
  st_crop(st_bbox(select.fires)) 
```

4. Select areas of public land that are 1000 m away from a road
```{r, eval=F}
sf_use_s2(FALSE) # turn off spherical geometry

hwys1000 <- hwys %>% st_buffer(dist=1000)
major1000 <- major %>% st_buffer(dist=1000)
local1000 <- local %>% st_buffer(dist=1000)
select.usfs.roads1000 <- select.usfs.roads %>% st_buffer(dist=1000)
roadz1000 <- st_union(hwys1000, major1000) %>% st_union(local1000) %>% st_union(select.usfs.roads1000)

hwys500 <- hwys %>% st_buffer(dist=500)
major500 <- major %>% st_buffer(dist=500)
local500 <- local %>% st_buffer(dist=500)
select.usfs.roads500 <- select.usfs.roads %>% st_buffer(dist=500)
roadz500 <- st_union(hwys500, major500) %>% st_union(local500) %>% st_union(select.usfs.roads500)

hwys250 <- hwys %>% st_buffer(dist=250)
major250 <- major %>% st_buffer(dist=250)
local250 <- local %>% st_buffer(dist=250)
select.usfs.roads250 <- select.usfs.roads %>% st_buffer(dist=250)
roadz250 <- st_union(hwys250, major250) %>% st_union(local250) %>% st_union(select.usfs.roads250)

hwys100 <- hwys %>% st_buffer(dist=100)
major100 <- major %>% st_buffer(dist=100)
local100 <- local %>% st_buffer(dist=100)
select.usfs.roads100 <- select.usfs.roads %>% st_buffer(dist=100)
roadz100 <- st_union(hwys100, major100) %>% st_union(local100) %>% st_union(select.usfs.roads100)

hwys50 <- hwys %>% st_buffer(dist=50)
major50 <- major %>% st_buffer(dist=50)
local50 <- local %>% st_buffer(dist=50)
select.usfs.roads50 <- select.usfs.roads %>% st_buffer(dist=50)
roadz50 <- st_union(hwys50, major50) %>% st_union(local50) %>% st_union(select.usfs.roads50)

hwys30 <- hwys %>% st_buffer(dist=30)
major30 <- major %>% st_buffer(dist=30)
local30 <- local %>% st_buffer(dist=30)
select.usfs.roads30 <- select.usfs.roads %>% st_buffer(dist=30)
roadz30<- st_union(hwys30, major30) %>% st_union(local30) %>% st_union(select.usfs.roads30)
```

5. Select areas of public land that are 500 m away from a trail and no more than 1 km up the trail
```{r, eval=F}
# identify trails that intersect roads
trail.id <- st_intersects(roadz30, trails)

trailz <- trails[trail.id[[1]],]
 
trailz <- trailz %>% st_buffer(dist=500) # buffer by 500 m
trailz2 <- st_intersection(trailz, roadz1000) # select part of trail within 1 km of a road
```

```{r, eval=F}
accessible <- st_union(trailz2, roadz250)
st_write(accessible, here("Data/Spatial/Sampling/accessible.shp"), append=F)

burned.public.accessible <- st_intersection(accessible, burned.public)  %>% st_union() %>% st_difference(roadz50)
st_write(burned.public.accessible, here("Data/Spatial/Sampling/burned-public-accessible.shp"), append=F)
```

6. Exclude area that was closed due to road damage or post-fire management activities
```{r, eval=F}
closed.roads <- st_read(here('Data/Spatial/Sampling/Closures/CameronPeak-RoadClosures.shp')) %>%
  st_transform(st_crs(burned.public.accessible)) %>% 
  st_buffer(dist=1000) %>% 
  st_union()
closed.trails <- st_read(here('Data/Spatial/Sampling/Closures/CameronPeak-TrailClosures.shp')) %>%
  st_transform(st_crs(burned.public.accessible)) %>% 
  st_buffer(dist=1000)%>% 
  st_union()
closed.roads.trails <- st_union(closed.roads, closed.trails)

mulch <- st_read(here('Data/Spatial/Sampling/Closures/CameronPeak-Mulch.shp')) %>%
  st_transform(st_crs(burned.public.accessible)) %>% 
  st_buffer(dist=50) %>% 
  st_union()

closed <- st_union(closed.roads.trails, mulch)
burned.public.accessible.closed <-  st_difference(burned.public.accessible, closed)
st_write(burned.public.accessible.closed, here("Data/Spatial/Sampling/burned-public-accessible-closed.shp"), append=F)
```

7. Identify areas >100 m and <1000 m away from pre-fire aspen patch
```{r, eval=F}
aspen <- read_stars(here("Data/Spatial/Aspen/SRM_ASPEN_CLASSIFIED_RF_10M_V1_0522.tif"))
bb <- st_transform(select.fires, st_crs(aspen))
aspen <- aspen %>% st_crop(st_bbox(bb))

aspen.p <- st_as_sf(aspen[1], as_points = FALSE, merge =TRUE) 

aspen.p1 <- aspen.p %>%
  filter(SRM_ASPEN_CLASSIFIED_RF_10M_V1_0522.tif==1) %>% 
  st_union()
st_write(aspen.p1, here("Data/Spatial/Aspen/aspen.shp"), append=F)

aspen.p1.sub <- st_transform(burned.public.accessible,  st_crs(aspen.p1)) %>% st_buffer(dist=1000) %>% st_union()
aspen.p1.sub <- st_intersection(aspen.p1.sub, aspen.p1) 
st_write(aspen.p1.sub , here("Data/Spatial/Sampling/aspen-sub.shp"), append=F)

aspen.p1.100 <-aspen.p1.sub %>% 
  st_buffer(dist=100) %>%
  st_union()
st_write(aspen.p1.100, here("Data/Spatial/Sampling/aspen-p1-sub-100.shp"), append=F)

aspen.p1.1000 <- aspen.p1.100 %>% 
  st_buffer(dist=900) %>%
  st_union()
st_write(aspen.p1.1000, here("Data/Spatial/Sampling/aspen-p1-sub-1000.shp"), append=F)

aspen.p1.100.1000 <- st_difference(aspen.p1.1000, aspen.p1.100) %>%
  st_union()
st_write(aspen.p1.100.1000, here("Data/Spatial/Sampling/aspen-p1-sub-100-1000.shp"), append=F)

burned.public.accessible.closed <- st_transform(burned.public.accessible.closed, st_crs(aspen))
burned.public.accessible.withaspen.100.1000 <- st_intersection(burned.public.accessible.closed,aspen.p1.100.1000 )
burned.public.accessible.withaspen.100.1000 <- st_as_sf(burned.public.accessible.withaspen.100.1000 ) %>% mutate(value=1)

st_write(burned.public.accessible.withaspen.100.1000 , here("Data/Spatial/Sampling/burnedpublic-accessible-aspen-p1-low-100-1000.shp"), append=F)
```

8. Create a regular sample so that points are all at least 250 m apart

```{r, eval=F}
# create regular sample
sampleareaz.grid <- as(make_grid(burned.public.accessible.withaspen.100.1000 , res=250), "Raster")
sampleareaz.grid  <- fasterize(burned.public.accessible.withaspen.100.1000, sampleareaz.grid , field="value")
sampleareaz.pts <- st_as_sf(raster::rasterToPoints(sampleareaz.grid, spatial=T))
```

9. Select points that burned at high severity
```{r, eval=F}
## Assign aet, cwd, elevation, watershed identity, and aspen suitability
aetr <-  read_stars(here("Data/Spatial/Climate/TraitBasedApproach_Rodman/AET and CWD/AET_1981-2010.tif"))
cwdr <-  read_stars(here("Data/Spatial/Climate/TraitBasedApproach_Rodman/AET and CWD/CWD_1981-2010.tif"))
elevr <- read_stars(here("Data/Spatial/Elev/LF2020_Elev_220_CONUS/Tif/LC20_Elev_220.tif"))
aspen.suitability <- read_stars(here("Data/Spatial/Aspen/AspenDistribution/Coops-AspenModeledFreq1950-2006.tif"))
watershed <- st_read(here("Data/Spatial/Watershed/WBDHU12/WBDHU12.shp"))

xseverity <- st_extract(fireseverity, st_transform(sampleareaz.pts, st_crs(fireseverity)))
xaetr <- st_extract(aetr, st_transform(sampleareaz.pts, st_crs(aetr)))
xcwdr <- st_extract(cwdr, st_transform(sampleareaz.pts, st_crs(cwdr)))
xelevr <- st_extract(elevr, st_transform(sampleareaz.pts, st_crs(elevr)))
xaspen.suitability <- st_extract(aspen.suitability, st_transform(sampleareaz.pts, st_crs(aspen.suitability)))
xwater <- st_join(st_transform(sampleareaz.pts, st_crs(watershed)), watershed)
xfire <- st_join(st_transform(sampleareaz.pts, st_crs(select.fires)), select.fires)
xseverity <- st_extract(fireseverity, st_transform(sampleareaz.pts, st_crs(fireseverity)))

sampleareaz.pts.latlon <- st_transform(sampleareaz.pts, CRS("+init=epsg:4269"))
sampleareaz.pts <- sampleareaz.pts %>% mutate(Lat = st_coordinates(sampleareaz.pts.latlon)[,2], Lon =st_coordinates(sampleareaz.pts.latlon)[,1], Easting=st_coordinates(sampleareaz.pts)[,1], Northing= st_coordinates(sampleareaz.pts)[,2], Fire = xfire$Incid_Name, Severity = xseverity$mtbs_CONUS_2020.tif, Watershed = xwater$Name, AET = xaetr$`AET_1981-2010.tif`, CWD=xcwdr$`CWD_1981-2010.tif`, Elev = xelevr$LC20_Elev_220.tif, Suitability=xaspen.suitability$`Coops-AspenModeledFreq1950-2006.tif`)

cp.pts <- sampleareaz.pts %>%  filter(Fire== "CAMERON PEAK")
cp.pts <- cp.pts %>%  filter(Severity==4)
cp.pts$ID <- row.names(cp.pts)
cp.pts$Name<-paste0(row.names(cp.pts), "-", cp.pts$Watershed)

#cp.pts025 <- cp.pts %>%  filter(Elev >quantile(cp.pts$Elev)[1] & Elev <quantile(cp.pts$Elev)[2]) %>% #sample_n(20) %>% mutate(Elev.class = "low")
#
#cp.pts2550 <- cp.pts %>%  filter(Elev >quantile(cp.pts$Elev)[2] & Elev <quantile(cp.pts$Elev)[3]) %>% #sample_n(20) %>% mutate(Elev.class = "low-moderate")
#
#cp.pts5075 <- cp.pts %>%  filter(Elev >quantile(cp.pts$Elev)[3] & Elev <quantile(cp.pts$Elev)[4])  %>% #sample_n(20) %>% mutate(Elev.class = "moderate-high")
#
#cp.pts75100 <- cp.pts %>%  filter(Elev >quantile(cp.pts$Elev)[4] & Elev <quantile(cp.pts$Elev)[5]) %>% sample_n(20) %>% mutate(Elev.class = "high")

#cp.pts <- rbind(rbind(rbind(cp.pts025, cp.pts2550), cp.pts5075), cp.pts75100)

  
st_write(cp.pts, here("Data/Spatial/Sampling/Sampleptz-CP.shp"), append=F)
write.csv(as.data.frame(cp.pts), here("Data/Spatial/Sampling/Sampleptz.csv"))
```


