---
title: "POTR"
author: "Sarah Carter"
date: "8/29/2022"
output: 
  bookdown::html_document2:
    code_folding: hide
    toc: TRUE
    toc_depth: 4
    toc_float: TRUE
    collapsed: FALSE
    theme: spacelab
    fig_caption: TRUE
    number_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
POTR <- read.csv("/Users/sarahcarter/Downloads/POTR Seedling Data - Sheet1.csv")
#str(POTR)
POTR_site <- read.csv("/Users/sarahcarter/Downloads/POTR SITE DATA - Sheet1.csv")
#str(POTR_site)
REGEN <- read.csv("/Users/sarahcarter/Downloads/ALL REGEN - Sheet1.csv")

#rename variables because they are confusing 
POTR$site.name <- POTR$SITE.NAME
POTR$site.Number <- POTR$SITE..
POTR$height <- POTR$Height..cm.
```

#merging site and seedling data
```{r}
compiled <- merge(POTR, POTR_site, by.x = "SITE..", by.y = "Site.Number")
#compiled
```

#packages
```{r}
#install.packages("ggplot2")
library(dplyr)
library(tidyr)
library(ggplot2)
```


```{r}
#dotplot height by cluster
ggplot(compiled,aes(x=site.name,y=height, col = height)) + geom_point() + 
    xlab("Site Cluster Name") + 
    ylab(expression(paste("Height cm"))) +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle(" Seedling Height by Cluster")
```

```{r}
#boxplot height by cluster
ggplot(POTR, aes(x=site.name, y=height)) + 
  geom_boxplot() +
  xlab("Site Cluster") +
  ylab(expression(paste("Height cm"))) +
  ggtitle("Seedling Height by Cluster")

```

```{r}
#histogram counts
count <- count(compiled, site.name)
count


#filter for seedling density
dense <- compiled %>% filter(SITE.NAME == "CAM"| SITE.NAME == "LONG"| SITE.NAME == "MONTY"| SITE.NAME == "RAWAH" | SITE.NAME == "SNOW")

#height by cluster for dense clusters 
ggplot(dense,aes(x=site.name,y=height, col = height)) + geom_point() + 
    xlab("Site Cluster Name") + 
    ylab(expression(paste("Height cm"))) +
    theme_bw(base_size = 16) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle(" Seedling Height by Cluster")

```

#Mean plots
```{r}
library("ggpubr")
ggline(dense, x = "site.name", y = "Height..cm.", color = "site.name", add = c("mean_se", "jitter"),  ylab = "Height (cm)", xlab = "Cluster", labcol = "CLuster")

```

#height means by cluster
```{r}
tapply(compiled$Height..cm., compiled$SITE.NAME, summary)
```


Another random interesting plot type:
```{r}
#install.packages("ggridges")
library(ggridges)
#ridgeplot
ggplot(POTR, aes(x = height, y = site.name, fill=site.name)) +
  geom_density_ridges() +
  xlab("Height cm") +
  ylab(expression(paste("Cluster Name"))) +
  ggtitle("Seedling Height Distribution")
```

#data filtering for tests
```{r}
library(dplyr)

#by cluster only
snow <- compiled %>% filter(SITE.NAME =="SNOW")
res <- compiled %>% filter(SITE.NAME =="RES")
rawah <- compiled %>% filter(SITE.NAME =="RAWAH")
monty <- compiled %>% filter(SITE.NAME =="MONTY")
long <- compiled %>% filter(SITE.NAME =="LONG")
lake <- compiled %>% filter(SITE.NAME =="LAKE")
fish <- compiled %>% filter(SITE.NAME =="FISH")
elkhorn <- compiled %>% filter(SITE.NAME =="ELKHORN")
cr69 <- compiled %>% filter(SITE.NAME =="CR69")
cam <- compiled %>% filter(SITE.NAME =="CAM")
blue <- compiled %>% filter(SITE.NAME =="BLUE")

#by substrate
ash <- compiled %>% filter(Substrate == "A"| Substrate == "A/B"| Substrate == "A/L"| Substrate == "A/M")
bryophyte <- compiled %>% filter(Substrate == "B"| Substrate == "B/A"| Substrate == "B/L"| Substrate == "B/M")
litter <- compiled %>% filter(Substrate == "L"| Substrate == "L/A"| Substrate == "L/B"| Substrate == "L/M")
mineral <- compiled %>% filter(Substrate == "M"| Substrate == "M/A"| Substrate == "M/B"| Substrate == "M/L")
rock <- compiled %>% filter(Substrate == "R")
wood <- compiled %>% filter(Substrate == "W")

#by small topo
cc <- compiled %>% filter(Small.Topo == "CC")
cv <- compiled %>% filter(Small.Topo == "CV")
f <- compiled %>% filter(Small.Topo == "F")
s <- compiled %>% filter(Small.Topo == "S")

#by large topo
CC <- compiled %>% filter(Large.Topo == "CC")
CV <- compiled %>% filter(Large.Topo == "CV")
F <- compiled %>% filter(Large.Topo == "F")
S <- compiled %>% filter(Large.Topo == "S")

#by elevation groups
high <- compiled %>% filter(Elevation > "3064")
moderate <- compiled %>% filter(between(Elevation, 2713,3064))
low <- compiled %>% filter(Elevation < "2713") 

#by large CWD
Lcwdp <- compiled %>% filter(Large.CWD == "1")
Lcwda <- compiled %>% filter(Large.CWD == "0")

#by small CWD
Scwdp <- compiled %>% filter(Small.CWD == "1")
Scwda <- compiled %>% filter(Small.CWD == "0")
```

#seedling count variable
```{r}
seedling_count_site <- count(compiled, compiled$SITE..)
seedling_count_cluster <- count(compiled, compiled$SITE.NAME)
```

#method of deciding relative site elevation groups
```{r}
#median(POTR_site$Elevation)
#min(POTR_site$Elevation)
#max(POTR_site$Elevation)
#summary(POTR_site$Elevation)
```



#substrate tallies and chi-sq
```{r}
count(ash)
count(bryophyte)
count(litter)
count(mineral)
count(rock)
count(wood)

substrate_counts = c(224, 256, 8, 55, 0, 4)
#chisq.test(substrate_counts, correct = F)

```

#topo
```{r}
cc
cv
s
f
CC
CV
S
F
```


#relationships between height and other variables
```{r}
#height and substrate
kruskal.test(Height..cm. ~ Substrate, data = compiled)

library(FSA)
dunnTest(Height..cm. ~ Substrate, data = compiled, method = "holm")

#height and cluster
kruskal.test(Height..cm. ~ SITE.NAME, data = compiled)

dunnTest(Height..cm. ~ SITE.NAME, data = compiled, method = "holm")

#height and elevation

kruskal.test(Height..cm. ~ Elevation, data = compiled)

dunnTest(Height..cm. ~ Elevation, data = compiled, method = "holm")

#small topo and height

kruskal.test(Height..cm. ~ Small.Topo, data = compiled)

dunnTest(Height..cm. ~ Small.Topo, data = compiled, method = "holm")

#large topo and height

kruskal.test(Height..cm. ~ Large.Topo, data = compiled)

dunnTest(Height..cm. ~ Large.Topo, data = compiled, method = "holm")


```

#Kruskall tests with seedling count
```{r}
#count and cluster
kruskal.test(seedling ~ SITE.NAME, data = compiled)

dunnTest(seedling ~ SITE.NAME, data = compiled, method = "holm")

#count and elevation

kruskal.test(seedling ~ Elevation, data = compiled)

dunnTest(seedling ~ Elevation, data = compiled, method = "holm")

#count and topo 

kruskal.test(seedling ~ Small.Topo, data = compiled)

dunnTest(seedling ~ Small.Topo, data = compiled, method = "holm")


#count topo and height

kruskal.test(seedling ~ Large.Topo, data = compiled)

dunnTest(seedling ~ Large.Topo, data = compiled, method = "holm")
```


#substrate
```{r}
ash
bryophyte
litter
mineral
rock
wood
```




#elevation hist
```{r}
low
moderate
high
hist(compiled$Elevation, breaks = 10, col = "light blue", main = "Elevation range of seedlings across all sites", xlab = "Elevation (m)", ylab = "Frequency")
```
#browse and CWD

```{r}
#by large CWD
Lcwdp 
Lcwda 

#by small CWD
Scwdp 
Scwda 
```

#all regen
```{r}
#REGEN
abla <- REGEN %>% filter(ABLA > "0")
pico <- REGEN %>% filter(PICO > "0")
pien <- REGEN %>% filter(PIEN > "0")
pipo <- REGEN %>% filter(PIPO > "0")
psme <- REGEN %>% filter(PSME > "0")

sucker_regen <- REGEN %>% filter(POTR..sucker. > "0")

abla
pico
pien
pipo
psme
sucker_regen
```

#other aspen
```{r}
sucker <- compiled %>% filter(Sucker.Dist. < "51")
parent <- compiled %>% filter(Distance.to.nearest.live.aspen < "51")
orphan <- compiled %>% filter(Distance.to.nearest.dead.aspen < "51")

parent
orphan

```
#browse
```{r}
browse <- compiled %>% filter(Browse == "1")
no_browse <- compiled %>% filter(Browse == "0")

browse
tapply(browse$Height..cm., browse$SITE.., summary)
```

```{r}
compiled
seedling_count_cluster
```

```{r}
TransectCount <- count(compiled, compiled$Transect)
```

#how to set up binomial linear mixed model??
```{r}
n <- 38
id <- seq(n)
transect <- 1:2
d <- expand.grid(id = id, transect=transect)
set.seed(1)
#sect <- sample(c("A", "B"), size = n, replace = TRUE)
cluster <- sample(c("Blue", "Cam", "CR69", "Elkhorn", "Fish", "Lake", "Long", "Monty", "Rawah", "Res", "Snow"), size = n, replace = TRUE)
d$cluster <- cluster[d$id]
#d$sect <- sect[d$id]
d <- d[order(d$id, d$transect),]
rownames(d) <- NULL
head(d, n = 76)
```

Probabilities of seedling presence:
```{r}
#Blue p = 
#Cam
#CR
#Elk
#Fish
#Lake
#Long
#Monty
#Rawah
#Res
#Snow



d$clusterint <- interaction(d$cluster)
probs <- c()
```

#Visualizing abnormality
```{r}
#hist(seedling_count_cluster$n)
#hist(seedling_count$n)
#hist(compiled$Height..cm.)
```





















